---
title: "R Notebook"
output: html_notebook
---

# Question 2

## Necessary library

Necessary library was imported into R.

```{r}
library(tidyverse)
library(readxl)
library(caret)
```

## Importing data

The data will first be imported into R, and observe the summary statistics of the data.

```{r}
data <- read_excel("Fish Weight-3Only-Random.xlsx")
attach(data)
summary(data)
str(data)
head(data)
```

## One-hot coding

Doing one-hot coding for Bream, Roach, and Perch.

```{r}
data <- data %>% 
  mutate("Bream-D" = ifelse(Species == "Bream",1,0),
         "Roach-D" = ifelse(Species == "Roach",1,0),)
head(data)
tail(data)
```

## Spliting data 80-20

The data is split into 80-20 for training-test sets for each `Species`.

```{r}
BreamTraining <- filter(data, Species == "Bream") %>% 
  head(dim(.)[1]*0.8)
BreamTest <- filter(data, Species == "Bream") %>% 
  tail(dim(.)[1]*0.2)

tableBream <- c(RawSize = dim(filter(data,Species == "Bream"))[1],
                TrainingSize = dim(BreamTraining)[1],
                TestSize = dim(BreamTest)[1])

RoachTraining <- filter(data, Species == "Roach") %>% 
  head(dim(.)[1]*0.8)
RoachTest <- filter(data, Species == "Roach") %>% 
  tail(dim(.)[1]*0.2)

tableRoach <- c(RawSize = dim(filter(data,Species == "Roach"))[1],
                TrainingSize = dim(RoachTraining)[1],
                TestSize = dim(RoachTest)[1])

PerchTraining <- filter(data, Species == "Perch") %>% 
  head(dim(.)[1]*0.8)
PerchTest <- filter(data, Species == "Perch") %>% 
  tail(dim(.)[1]*0.2)

tablePerch <- c(RawSize = dim(filter(data,Species == "Perch"))[1],
                TrainingSize = dim(PerchTraining)[1],
                TestSize = dim(PerchTest)[1])

splitSize <- rbind(tableBream, tableRoach, tablePerch)  
row.names(splitSize) <- c("Bream","Roach","Perch")
splitSize

TrainingSet <- rbind(BreamTraining, RoachTraining, PerchTraining)
TestSet <- rbind(BreamTest, RoachTest, PerchTest)
```

## Linear regression

The training data will then trained to fit into the linear model as follow:

$Weight = \beta_0 + \beta_1\cdot Height+\beta_2\cdot Width + \beta_3\cdot \mbox{Bream-D} + \beta_4\cdot\mbox{Roach-D}$

```{r}
model1 <- train(Weight ~ Height + Width + `Bream-D` + `Roach-D`,
                data = TrainingSet, method = "lm")

summary(model1)
print(model1)
```


