---
title: "R Notebook"
output: html_notebook
---

# Question 3

## Import library

```{r}
library(tidyverse)
library(readxl)
library(caret)
library(PerformanceAnalytics)
library(car)
library(factoextra)
library(pls)
```


## Data Import

First, let us import the data.

```{r}
data <- read_csv("CarBig.csv")
```

## Data Exploration

We will first have a glimpse on the data

```{r}
dim(data)
summary(data)
head(data)
str(data)
```

This data set has 13 variables. First, we can convert the categorical data to factors for easier manipulation later and determine the variables that are unnecessary.

```{r}
data$cyl4 <- as.factor(data$cyl4)
data$Mfg <- as.factor(data$Mfg)
data$Origin <- as.factor(data$Origin)
data$org <- as.factor(data$org)
data$when <- as.factor(data$when)
data$Model_Year <- as.factor(data$Model_Year)
```

Then we re-inspect the data.

```{r}
str(data)
```

We can see that `cyl4` is just another presentation of variable `Cylinders` stating whether its 4 cylinder, hence we do not need the variable `cyl4`. Next, we see that `org` and `Origin` are virtually describing the same variable. Hence, we keep the variable `org` instead of `Origin` as we want the European countries to be represented under the factor `Europe`. We will drop `when` variable too as it only describe when in the year the car is released.

```{r}
data <- select(data, -c(cyl4, Origin, when))
str(data)
```

We will do some quick exploratory plotting on the numerical data.

```{r}
select(data, Acceleration, Displacement, Horsepower, MPG, Weight, Cylinders) %>% 
  chart.Correlation(histogram = TRUE, pch = 19)
```

We can see that the `Acceleration` is more or less normally distributed while `Displacement`, `Horsepower`, `MPG`, and `Weight` are roughly log-normally distributed. All the correlation marked with 3 stars has a $p<=0.001$. Since are variables are highly correlated (except MPG-Acceleration and MPG-Weight pairs which are moderately positively correlated) with each other, we will do an VIF test to see if there are any multicollinearity problem occurs within the predictors later.

We will now do some plotting to see how the categorical data fits in the scatter plot.

```{r}
ggplot(data)+
  geom_point(aes(x = Acceleration, y = MPG, color = Weight, size = Horsepower))

```

From the plot, we can see that usually cars with higher `Weight` have lower `MPG` and `Acceleration` probably due to more fuel and force is needed to carry the higher weight. 

```{r}
ggplot(data)+
  geom_point(aes(x = Horsepower, y = MPG, color = Cylinders))
```

Now from this plot, we can see that cars with more `Cylinders` tends to have more `Horsepower`, which makes sense and lower `MPG`. This might due to the increase in weight causing more insufficient consumption of fuel.

Next, we are interested to see how `MPG`, `Horsepower`, `Weight` and `org` is correlated to each other. We are interested to see country from which origin has the highest miles-per-gallon and horsepower.

```{r}
ggplot(data)+
  geom_point(aes(x = Horsepower, y = MPG, size = Weight, color = org))
```

We can see that most cars from Europe and Japan has smaller weight, greater `MPG` and lesser `Horsepower`. The small weight inferred that the car is more compact, less cylinders in the engine which leads to lower `Horsepower` but greater performance in terns of fuel saving `MPG`.

## Data transformation while doing linear fitting

```{r}
select(data, Acceleration, Displacement, Horsepower, MPG, Weight, Cylinders) %>% 
  chart.Correlation(histogram = TRUE, pch = 19)
```

Observing from this graph and having `MPG` as our output, we can do some transformation to the features. We see that `Displacement`, `Horsepower`, `Weight` and `Cylinders` are inversely proportional to `MPG`, hence we will do a $1/x$ transformation to these features. However, due to the nature that `Displacement`, `Horsepower`, `Weight` and `Cylinders` are highly correlated, we expect that only one is needed in the end. We will do ANOVA test to compare model fitting. The categorical variables `Model`, `Mfg` and `Model_Year` of the car are not needed in the linear regression as they are already shown to reflect their effect through the numerical variable as shown in previous exploratory analysis.

### Data radomizing

First, we will shuffle our data so that the data is not sorted in any patterns.

```{r}
set.seed(100)
data <- data %>% 
  mutate(rand = runif(dim(data)[1])) %>% 
  arrange(rand)

head(data)
```

Then we subset out the features that are needed. Here we subset the columns `Acceleration`, `Cylinders`, `Displacement`, `Horsepower`, `MPG`, and `Weight`.

```{r}
data <- data %>% 
  select(-c(Mfg, Model, Model_Year, org, rand)) %>% # unselect the stated features
  mutate(invCylinders = 1/Cylinders,
         invHorsepower = 1/Horsepower,
         invDisplacement = 1/Displacement,
         invWeight = 1/Weight)
```

### Training and test set

Next, we split our data into 80% training set and 20% test set.

```{r}
dim(data)
```

As we can see, there are 406 observations. Hence, we will select top 325(~80%) out of the randomized data as training set and the remaining (81) as test set.

```{r}
trainingSet <- head(data, 325)
testSet <- tail(data, 81)
```

### Linear fitting

Now, we will do our linear fitting with all the variables with the $1/x$ transformation applied to `Horsepower`, `Displacement`, and `Weight`. We will be omitting observations with NaN values.

$\hat{MPG_1}=\hat{\beta_0}+\hat{\beta_1}\cdot Acceleration + \hat{\beta_2}\cdot \frac{1}{Cylinders} +\hat{\beta_3}\cdot\frac{1}{Displacement} +\hat{\beta_4}\cdot \frac{1}{Horsepower} + \hat{\beta_5}\cdot \frac{1}{Weight}$

```{r}
model1 <- train(MPG ~ Acceleration + invCylinders + invDisplacement +
                  invHorsepower + invWeight,
                data = trainingSet,
                na.action = na.omit,
                method = "lm")

model1
summary(model1)
```

Now we will do VIF test and with p-values of the t-statistics to decide which features to drop.

```{r}
vif(model1$finalModel)
```

We can see that from the VIF test, `invCylinders` and `invWeight` are highly correlated to each other as they has high VIF value. Besides, `invDisplacement` also has a non-significant p-value for the t-statistics, hence we will drop the tern `invDisplacement`.

$\hat{MPG_2}=\hat{\beta_0}+\hat{\beta_1}\cdot Acceleration + \hat{\beta_2}\cdot \frac{1}{Cylinders} + \hat{\beta_3}\cdot \frac{1}{Horsepower} + \hat{\beta_4}\cdot \frac{1}{Weight}$

```{r}
model2 <- train(MPG ~ Acceleration + invCylinders +
                  invHorsepower + invWeight,
                data = trainingSet,
                na.action = na.omit,
                method = "lm")

model2
summary(model2)
```

We see that the model performance has increased from $R^2_{adj}=0.7457$ to $R^2_{adj}=0.7465$. We will do an ANOVA test to see if both model1 and model2 are performing equally well.

```{r}
anova(model1$finalModel, model2$finalModel)
```

The ANOVA test has a $p=0.9941$ implying that we cannot reject the null hypothesis that both model are performing equally well. We do not have sufficient evidence that model2 is performing better than model1 with the training data.

Now, we run again VIF test for model2.

```{r}
vif(model2$finalModel)
```

Through the VIF test, we can see that `invHorsepower` and `invWeight` are highly correlated of value greater than 5. We will remove the term `invWeight` as it has a higher p-values for t-statistics to avoid multi-collinearity problem.

$\hat{MPG_3}=\hat{\beta_0}+\hat{\beta_1}\cdot Acceleration + \hat{\beta_2}\cdot \frac{1}{Cylinders} + \hat{\beta_3}\cdot \frac{1}{Horsepower}$

```{r}
model3 <- train(MPG ~ Acceleration + invCylinders + invHorsepower,
                data = trainingSet,
                na.action = na.omit,
                method = "lm")

model3
summary(model3)
```

Now we see that the model3 has a slightly smaller $R^2_{adj}=0.7353$. We will run an ANOVA test to see if model3 performs equally well to model2.

```{r}
anova(model2$finalModel, model3$finalModel)
```

The ANOVA test has a $p=0.0001629<0.005$ implying that we have sufficient evidence to reject the null hypothesis that model2 and model3 performs equally well. 

However, one problem to note that there might be a multi-collinearity problem in model2 observing from the VIF test, which might in turn affects the validity of p-values for the ANOVA test. The slight drop in $R^2_{adj}$ worth the reduction in the multi-collinearity problem. Hence, we will keep model3 as a better model than model2.

Since `Horsepower` is non-linear, we will try to fit `Horsepower` to the 2nd order polynomial instead of $1/x$.

$\hat{MPG_4}=\hat{\beta_0}+\hat{\beta_1}\cdot Acceleration + \hat{\beta_2}\cdot \frac{1}{Cylinders} + \hat{\beta_3}\cdot Horsepower + \hat{\beta_4}\cdot Horsepower^2$

```{r}
model4 <- trainingSet %>% 
  mutate(Horsepower2 = Horsepower^2) %>% 
  train(MPG ~  Acceleration + invCylinders + Horsepower + Horsepower2,
                data = .,
                na.action = na.omit,
                method = "lm")

model4
summary(model4)
```

We can see that it has a slightly lower $p=0.7342$ compared to model3. 

So far model2 has the higher $R^2_{adj}$ but it suffers multi-collinearity problem. Hence, the next model with highest training $R^2_{adj}$ is model3.

## Regress on test set

Now, we will apply our model3 and model4 on test set and observe their performance. Model1 and model2 are not being considered due to multi-collienarity problem. First, we will create the metric reporting function that reports `RMSE`, `R2` and `R2_adj`.

```{r}
metric <- function(raw, predict, nPredictor, dp = 5){
  n <- length(predict)
  SSE <- sum((raw - predict) ^ 2)
  SST <- sum((raw - (sum(raw)/n)) ^ 2)
  R2 <- 1 - SSE/SST
  R2adj <- 1 - (((1 - R2)*(n - 1))/(n - nPredictor - 1))
  RMSE <- sqrt(SSE/n)
  table <- data.frame(cbind(n, round(SSE,dp), round(SST,dp), round(R2,dp),
             round(R2adj,dp), round(RMSE,dp)))
  colnames(table) <- c("n","SSE","SST","R2","R2adj","RMSE")
  return(table)
}
```

Now we will apply all 4 models on the test data.

```{r}
predictModel3 <- na.omit(testSet) %>% 
  predict(model3, .)

predictModel4 <- na.omit(testSet) %>% 
  mutate(Horsepower2 = Horsepower^2) %>% 
  predict(model4, .)

metricModel3 <- metric(na.omit(testSet)$MPG, predictModel3, length(model3$finalModel$coefficients)-1)
metricModel4 <- metric(na.omit(testSet)$MPG, predictModel4, length(model4$finalModel$coefficients)-1)
rbind(metricModel3, metricModel4)
```

We can see that both model3 and model4 performs roughly equally well according to the $R^2_{adj}$ generated.

## Principal Component Analysis

We will be doing PCR. First, we will get the Training Set and Test set again, and select only relevant variables.

```{r}
testSet <- na.omit(testSet)
trainingSet <- na.omit(trainingSet)
```

Then we will perform PCA

```{r}
pcr1 <- trainingSet %>% 
  select(Acceleration, Cylinders, Displacement, Horsepower, MPG, Weight) %>% 
  pcr(MPG ~., data = ., scale = TRUE) 
                 
summary(pcr1)
```

According to PCA, with 95% variance, only 2 PCA component is needed, with a cumulative variance percent of 95.25%.

```{r}
pcrPredict1 <- predict(pcr1, testSet, ncomp = 2)
metric(testSet$MPG, pcrPredict1, length(pcr1$coefficients)-1) 
```

We can see that PCR without the transformed variables does not perform as well compared to the linear regression with transformed variable with that significantly smaller $R^2_{adj}$ value.

### PCR with transformed variable

Since PCR does not work well if the variables are not linear, we will try again PCR with the linearly transformed variables: `invWeight`, `invCylinders`, `invHorsepower`, and `invDisplacement`.

```{r}
pcr2 <- trainingSet %>% 
  select(Acceleration, invCylinders, invDisplacement, invHorsepower, MPG, invWeight) %>% 
  pcr(MPG ~., data = ., scale = TRUE) 
                 
summary(pcr2)
```

With the inverse variables, we can see that with a threshold variance of 95%, 3 components must be chosen with a accumulated 97.49% of variance explained.

```{r}
pcrPredict2 <- predict(pcr2, testSet, ncomp = 3)
metric(testSet$MPG, pcrPredict2, length(pcr2$coefficients)-1) 
```

From the metric generated, we can see that PCR with transformed variable performed better than PCR without transformed variable. This is probably due to the fact that the transformed variables are more closed to linear variable and the PCA algorithm we work with works on linear variable.























